Fractional Types

### 2016-12-19 typespp.tex
adjust introduction to fit the paper.

In Homotopy Type Theory (HoTT)~\cite{hottbook}, types have the
structure of \emph{weak $\omega$-groupoids}. As a first approximation,
we can think of such structures as sets with points (objects) and
paths (equivalences) between the points and higher paths between these
paths and so on. Here are two simple but non-trivial examples:

@@

Both groupoids involve some notion of ``division'' that can be captured at the
level of types using a syntactic notion of \emph{fractional types}. Our aim is
to explore this notion of types and its associated operational semantics. The
remainder of the paper is organized as follows. We start by reviewing the
necessary background consisting of the language $\Pi$ for programming with
isomorphisms or equivalences in a reversible information-preserving way. Sec.~4
explains the main novel semantic ideas of using $\Pi$ programs to generate
non-trivial groupoids with fractional cardinality. Sec.~5 translates the
semantic ideas into an extension of $\Pi$ with new type constructors denoting
non-trivial groupoids and new programs that manipulate such types. Sec.~6
presents the operational semantics of the extended language. The last section
puts our work in perspective and concludes.

@@

 [\order{c_2}.c_2^k]}}

>>>>

In modern treatments of type theory, types have the
structure of \emph{weak $\omega$-groupoids}. As a first approximation,
we can think of such structures as sets with points (objects) and
paths (equivalences) between the points and higher paths between these
paths and so on. Here are two simple but non-trivial examples:

@@

Both groupoids involve some notion of (semantic) ``equivalence'', which we
would like to capture as first-class entities in a programming language.
In other words, we would like to have some syntactic notion of a type, whose
denotation would be a groupoid with fractional cardinality.  In other words,
a syntactic notion of \emph{fractional types}. Our aim then is to create a
language which possesses ``fractional types'', explore this notion of types,
and its associated operational semantics.

The remainder of the paper is organized as follows. We start by reviewing
necessary background material, consisting of the language $\Pi$ for programming
with isomorphisms or equivalences in a reversible information-preserving way.
Sec.~4 explains the main novel semantic ideas of using $\Pi$ programs to
generate non-trivial groupoids with fractional cardinality, as well as
translating the semantic ideas into an extension of $\Pi$ with new type
constructors denoting non-trivial groupoids and new programs that manipulate
such types.  Sec.~5 presents some example programs.  The last section puts our
work in perspective and concludes.

@@

 [\order{c_2},c_2^k]}}
{\jdg{}{}{(\iorder{\alpha})_k :
 [\iorder{c_1},1/c_1^k]  \leftrightblackspoon
 [\iorder{c_2},1/c_2^k]}}
